
---
title: 'PhD: CSE9099c - Machine Learning Algorithms'
author: "Jisha Joseph 1845"
date: "July-27-2018"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---
# Set up the environement
```{r}
rm(list=ls(all = T))
setwd("C:/Users/Hp/Desktop/INSOFE/Jisha Joseph_PhD")
```
# List & Load all required libraries here
```{r message = FALSE}
library(DMwR)
library(DataExplorer)
library(caret)
library(dplyr)
library(plyr)
library(ROCR)
library(corrplot)
library(class)
library(ada)
library(xgboost)
library(randomForest)
library(rpart)
library(C50)
library(e1071)
library(tidyverse)
library(lubridate)
library(tidyr)
library(sqldf)
library(ggplot2)
```

# Read & Understand the data
```{r}
train_data <- read.csv("Train.csv",header=TRUE)
test_data <- read.csv("Test.csv",header=TRUE)
```
## Keep an Original Copy of Train and test before doing any further processing
```{r}
train_data_orig<-train_data
test_data_orig <- test_data
```
## Get a feel of the data and get Insights
```{r}
head(train_data)
head(test_data)
tail(train_data)
tail(test_data)
```
## Structure & summary of the Train and Test data
```{r}
str(train_data)
summary(train_data)
summary(test_data) 
str(test_data)
```
# Target Attribute Derivation
## Convert ActualArrivalTimeStamp to POSIX type using Lubridate Library
```{r}
##Train data
train_data$ActualArrivalTSnew<-train_data[['ActualArrivalTimeStamp']]
train_data$ActualArrivalTSnew<-dmy_hm(train_data$ActualArrivalTSnew)
str(train_data)
head(train_data)
```
## Drop ActualArrivalTimeStamp as a new column has been created from it
```{r}
train_data$ActualArrivalTimeStamp=NULL
```
## Convert ScheduledArrTime to HH:MM format
```{r}
##Train Data
train_data$ScheduledArrTime<-sprintf("%04d",train_data$ScheduledArrTime)
format(strptime(train_data$ScheduledArrTime, format="%H%M"), format = "%H:%M")

##Test Data

test_data$ScheduledArrTime<-sprintf("%04d",test_data$ScheduledArrTime)
format(strptime(test_data$ScheduledArrTime, format="%H%M"), format = "%H:%M")
```
## Merge the Expected arrival into the same format of Actual Arrival
```{r}

##Train Data
train_data<-unite(train_data,"ExpectedArrival",c("Year","Month","DayofMonth","ScheduledArrTime"),sep = '/',remove = TRUE)
train_data$ExpectedArrival<-ymd_hm(train_data$ExpectedArrival)

##Test Data
test_data<-unite(test_data,"ExpectedArrival",c("Year","Month","DayofMonth","ScheduledArrTime"),sep = '/',remove = TRUE)
test_data$ExpectedArrival<-ymd_hm(test_data$ExpectedArrival)
```
## Calculate the difference in Minutes
```{r}
difference <- difftime(train_data$ActualArrivalTSnew,train_data$ExpectedArrival, units='mins')
#difference
```

## Check if the difference is greater than 15 mins, then Create the Target attribute-FlightDelayStatus as 'Yes' else 'No'
```{r}
FlightDelayStatus<-ifelse(difference > 15,'Yes','No')
#FlightDelayStatus
```
## Add the FlightDelayStatus to the Train dataframe
```{r}
train_data<-cbind(train_data,difference)
train_data<-cbind(train_data,FlightDelayStatus)
str(train_data)
str(test_data)
```
## Convert attributes to required data types
```{r}
## Train Data

train_data$DayOfWeek <- as.factor(as.character(train_data$DayOfWeek))

train_data$ScheduledDepTime<-as.numeric(train_data$ScheduledDepTime)
train_data$ScheduledDepTime<-sprintf("%04d",train_data$ScheduledDepTime)
format(strptime(train_data$ScheduledDepTime, format="%H%M"), format = "%H:%M")
train_data$difference<-as.numeric(train_data$difference)

## Test data

test_data$DayOfWeek <- as.factor(as.character(test_data$DayOfWeek))
test_data$ScheduledDepTime<-as.numeric(test_data$ScheduledDepTime)
test_data$ScheduledDepTime<-sprintf("%04d",test_data$ScheduledDepTime)
format(strptime(test_data$ScheduledDepTime, format="%H%M"), format = "%H:%M")

```

#Check Uniquesness of Flightnumber
```{r}
unique(train_data$FlightNumber)
unique(test_data$FlightNumber)
```
## Observations of Test and Train
*Remove Difference from Train in case of Classification/Keep in case of regression to derive Target as the same way as of Train
*DayOfWeek has near zero variance from the plot.Hence remove.
*Levels for Origin and Destination to be added to Test data
```{r}
train_data_bkp<-train_data
delete <- c("difference","DayOfWeek")
train_data<-train_data[,!(colnames(train_data) %in% delete),drop=FALSE]
test_data$DayOfWeek=NULL
```
## Univariate and Bivariate Analysis
* Univariate Analysis; Check if there are Outliers
```{r}
boxplot(train_data$ScheduledTravelTime)
```
* Bar Plot to check Kurtosis and Skewness
```{r}
ggplot(train_data,aes(x=FlightDelayStatus)) + geom_bar()
```
* Target is imbalanced as from the above plot.Hence have to balance the target
* Bivariate Analysis
```{r}
ggplot(train_data_orig,aes(x=train_data_orig$Distance,
           y=train_data_orig$ScheduledTravelTime)) + 
geom_point(size=2)
```

## Split Categorical and Numerical Attributes
```{r}
num_attr<-c("ScheduledDepTime","ScheduledArrTime","ScheduledTravelTime","Distance")
cat_attr<-setdiff(colnames(train_data_orig),num_attr)
cat_attr
```
## Check Correlation Plot
```{r}
num_attr_df<-data.frame(train_data_orig[,num_attr])
#is.numeric((train_data_orig[,"FlightNumber"]))
cor_num_attr_df<-cor(num_attr_df)
#corrplot(cor_num_attr_df, method = "square")
corrplot(cor_num_attr_df, method = "number")
```
## Remove Distance as it is highly correlated(0.98) with ScheduledTravelTime
## Remove ScheduledDepTime as it is highly correlated with ScheduledTravelTime
```{r}
train_data$Distance=NULL
test_data$Distance=NULL
train_data$ScheduledDepTime=NULL
test_data$ScheduledDepTime=NULL
```
# ###################Start of Weather Data Processing##########################
##Read AllStationsData_PHD as a dataframe
```{r}
AllStations<-read.table("AllStationsData_PHD.txt",sep="|", header=TRUE)
```
## Understand the data
```{r}
str(AllStations)
dim(AllStations)
```
## Split Categorical and Numerical Attributes for AllStations Data
```{r}
cat_attr_alls<-c("AirportID","TimeZone")
num_attr_alls<-setdiff(colnames(AllStations),cat_attr_alls)
num_attr_alls
```
## Check Correlation Plot
```{r}
num_attr_dfalls<-data.frame(AllStations[,num_attr_alls])
cor_num_attr_dfalls<-cor(num_attr_dfalls)
#corrplot(cor_num_attr_df, method = "square")
corrplot(cor_num_attr_dfalls, method = "number")
```
## Remove correlated attributes from AllStations Data before merging with Train Data
* Remove BarometerHeight as it is fully correlated with GroundHeight
* BarometerHeight and StationHeight are highly correlated.(0.84)
* Remove StationHeight as it is highly correlated with GroundHeight(0.84)
```{r}
AllStations$BarometerHeight=NULL
AllStations$StationHeight=NULL
```
## Missing Values Check and Imputation in AllStations
```{r}
sort(colSums(is.na(AllStations))/nrow(AllStations)*100, decreasing = T)
```
## Split Categorical and Numerical Attributes for AllStations Data after removal
```{r}
cat_attr_alls<-c("AirportID","TimeZone")
num_attr_alls<-setdiff(colnames(AllStations),cat_attr_alls)
num_attr_alls
```

## Check for the Origin and Destination Levels mismatched in Train and test
* Mismatched levels to be added
```{r}
unique(train_data$Origin[!(train_data$Origin %in% test_data$Origin)])
```
## Use sqldf for joining AllStations and Train data,Test Data;deleted AirportID
```{r}
train_allst<-sqldf("SELECT train_data.*,AllStations.* FROM train_data
LEFT OUTER JOIN AllStations ON AllStations.AirportID = train_data.Destination")
train_allst<-train_allst[,-6]  #-->Actual ActualArrivalTSnew
train_allst<-train_allst[,-8]  #-->AirportID

test_allst<-sqldf("SELECT test_data.*,AllStations.* FROM test_data
LEFT OUTER JOIN AllStations ON AllStations.AirportID = test_data.Destination")
test_allst<-test_allst[,-8]
```
#  Hourly and hpd processing
## Read 2004Hourly and HPD file to see the data
```{r}
df200401hrly<-read.table("200401hourly.txt",sep=",", header=TRUE)
df200401hpd<-read.table("200401hpd.txt",sep = ",",header = TRUE)
df200403hrly<-read.table("200403hourly.txt",sep=",", header=TRUE)
df200403hpd<-read.table("200403hpd.txt",sep = ",",header = TRUE)
df200405hrly<-read.table("200405hourly.txt",sep=",", header=TRUE)
df200405hpd<-read.table("200405hpd.txt",sep = ",",header = TRUE)
df200407hrly<-read.table("200407hourly.txt",sep=",", header=TRUE)
df200407hpd<-read.table("200407hpd.txt",sep = ",",header = TRUE)
df200409hrly<-read.table("200409hourly.txt",sep=",", header=TRUE)
df200409hpd<-read.table("200409hpd.txt",sep = ",",header = TRUE)
df200411hrly<-read.table("200411hourly.txt",sep=",", header=TRUE)
df200411hpd<-read.table("200411hpd.txt",sep = ",",header = TRUE)
```
## Merge all the hpd and hourly rows into one Dataframe
```{r}
df2004hrly<-rbind(df200401hrly,df200403hrly,df200405hrly,df200407hrly,df200409hrly,df200411hrly)
rm(df200403hrly,df200405hrly,df200407hrly,df200409hrly,df200411hrly)
df2004hpd<-rbind(df200401hpd,df200403hpd,df200405hpd,df200407hpd,df200409hpd,df200411hpd)
rm(df200403hpd,df200405hpd,df200407hpd,df200409hpd,df200411hpd)
#rm(df200401hpd) 
#rm(df200401hrly)
```
## Structure of HPD and Hourly
```{r}
str(df2004hpd)
str(df2004hrly)
```
## Missing Values Check and Imputation
```{r}
sort(colSums(is.na(df2004hrly))/nrow(df2004hrly)*100, decreasing = T)
sort(colSums(is.na(df2004hpd))/nrow(df2004hpd)*100, decreasing = T)
```
## Let's use Central imputation from DmWR package to impute missing values
```{r}
df2004hrly <- centralImputation(df2004hrly)
df2004hpd <- centralImputation(df2004hpd)
#df200401hpd<-centralImputation(df200401hpd) -->Testing Purpose
```
## Check if NA values are filled
```{r}
colSums(is.na(df2004hpd))
colSums(is.na(df2004hrly))
```

## Process hpd data 2004
## Convert and Merge the Timestamp in 2004hpd to Same as train_allst
## Convert ScheduledArrTime to HH:MM format
```{r}
##HPD Train Data
# Convert Time to Numeric and then to HH:MM format
df2004hpd$Time<-as.numeric(df2004hpd$Time)
df2004hpd$Time<-sprintf("%04d",df2004hpd$Time)
format(strptime(df2004hpd$Time, format="%H%M"), format = "%H:%M")
str(df2004hpd)
```

## Merge date and Time in hpd
```{r}
##hpd Train Data
df2004hpd<-unite(df2004hpd,"Timestamp",c("YearMonthDay","Time"),remove = TRUE)
df2004hpd$Timestamp<-ymd_hm(df2004hpd$Timestamp)
str(df2004hpd)
df2004hpdb<-df2004hpd

```

## Aggregate Time into 6-hour-wide intervals for 2004hpd and Train data
```{r}
## Train Data
#############
# Aggregate each Timestamp into corresponding hour
traints<-cut.POSIXt(train_allst$ExpectedArrival, breaks = "6 hours")
traints
traindf<-as.data.frame.Date(traints)
str(traindf)
traindf$traints<-as.POSIXct(traindf$traints)
traindf$traints<-traindf$traints - 3600
c<-cbind(train_allst,traindf)
train_allst<-c[,-2]  #-->Remove Actual Timestamp
colnames(train_allst)[11]<-"ExpectedArrival"

# HPD data
##########

df2004hpdts<-cut.POSIXt(df2004hpd$Timestamp, breaks = "6 hours")
b<-as.data.frame.Date(df2004hpdts)
str(b)
c<-cbind(df2004hpd,b)
df2004hpd<-c[,-2]  ##-->Remove Actual Time
colnames(df2004hpd)[3]<-"Timestamp"


df2004hpd<-aggregate(HourlyPrecip~WeatherStationID+Timestamp,data=df2004hpd,FUN=function(df2004hpd) mean(df2004hpd))


```
## Use merge for joining train_allst with hpd data
```{r}
trallst_hpd<-merge(x=train_allst,y=df2004hpd,by.x=c("WeatherStationID","ExpectedArrival"),by.y=c("WeatherStationID","Timestamp"),all.x=TRUE)

```

## Process hourly data 2004
## Split Categorical and Numerical Attributes for Hourly Data
```{r}
str(df2004hrly)
cat_attr_alls<-c("SkyConditions","Visibility","WindDirection","WindSpeed")
num_attr_alls<-setdiff(colnames(df2004hrly),cat_attr_alls)
num_attr_alls
```

## Check Correlation Plot for df2004hrly
```{r}
num_attr_dfalls<-data.frame(df2004hrly[,num_attr_alls])
cor_num_attr_dfalls<-cor(num_attr_dfalls)
#corrplot(cor_num_attr_df, method = "square")
corrplot(cor_num_attr_dfalls, method = "number")
```
## Remove correlated attributes
* Remove DewPointTemp as it is highly correlated(0.81) to DBT
* Remove WindDirection as it is highly correlated to Windspeed (p-value-2.2e-16)
* Remove Skyconditions as it is highly correlated to Visibility(p-value-2.2e-16)

```{r}
df2004hrly$DewPointTemp=NULL
df2004hrly$WindDirection=NULL
df2004hrly$SkyConditions=NULL
```

## Convert and Merge the Timestamp in 2004hourly to Same as trallst_hpd
## Convert ScheduledArrTime to HH:MM format
```{r}
##Hourly Train Data
# Convert Time to Numeric and then to HH:MM format
df2004hrlyb<-df2004hrly
df2004hrly$Time<-as.numeric(df2004hrly$Time)
df2004hrly$Time<-sprintf("%04d",df2004hrly$Time)
format(strptime(df2004hrly$Time, format="%H%M"), format = "%H:%M")
str(df2004hrly)
```

## Merge date and Time in hourly
```{r}
##hourly Train Data
df2004hrlyb<-df2004hrly
df2004hrly<-unite(df2004hrly,"Timestamp",c("YearMonthDay","Time"),remove = TRUE)
df2004hrly$Timestamp<-ymd_hm(df2004hrly$Timestamp)
str(df2004hrly)
```

## Aggregate Time into 6-hour-wide intervals for 2004hourly data
```{r}
# Hourly data
#############
# Aggregate each Timestamp into 6 hour window
df2004hrlyts<-cut.POSIXt(df2004hrly$Timestamp, breaks = "6 hours")
b<-as.data.frame.Date(df2004hrlyts)
str(b)
c<-cbind(df2004hrly,b)
df2004hrly<-c[,-2]
colnames(df2004hrly)[8]<-"Timestamp"

#Calculate the mean of HourlyPrecip grouped on WeatherStationID+Timestamp 
df2004hrly<-aggregate(cbind(Visibility,DBT,RelativeHumidityPercent,WindSpeed,WindGustValue,StationPressure)~WeatherStationID+Timestamp,data=df2004hrly,FUN=function(df2004hrly) mean(df2004hrly))

```

## Merge hourly data with trallst_hpd
```{r}
trallsthpd_hrly<-merge(x=trallst_hpd,y=df2004hrly,by.x=c("WeatherStationID","ExpectedArrival"),by.y=c("WeatherStationID","Timestamp"),all.x=TRUE)
```
## Check for NA values in the final trallsthpd_hrly dataframe
## Missing Values Check and Imputation
```{r}
sort(colSums(is.na(trallsthpd_hrly))/nrow(trallsthpd_hrly)*100, decreasing = T)
```
## Let's use Central imputation from DmWR package to impute missing values
```{r}
trallsthpd_hrly <- centralImputation(trallsthpd_hrly)
```
## Check if NA values are filled
```{r}
colSums(is.na(trallsthpd_hrly))
```
# Test hpd and hourly Processing
## Read 2005Hourly and HPD file to see the data
## Process 2005hpd data for test
```{r}
df200503hrly<-read.table("200503hourly.txt",sep=",", header=TRUE)
df200503hpd<-read.table("200503hpd.txt",sep = ",",header = TRUE)
df200507hrly<-read.table("200507hourly.txt",sep=",", header=TRUE)
df200507hpd<-read.table("200507hpd.txt",sep = ",",header = TRUE)
df200509hrly<-read.table("200509hourly.txt",sep=",", header=TRUE)
df200509hpd<-read.table("200509hpd.txt",sep = ",",header = TRUE)
df200511hrly<-read.table("200511hourly.txt",sep=",", header=TRUE)
df200511hpd<-read.table("200511hpd.txt",sep = ",",header = TRUE)
```
## Merge all the 2005hpd and hourly rows into one Dataframe for test
```{r}
df2005hrly<-rbind(df200503hrly,df200507hrly,df200509hrly,df200511hrly)
rm(df200503hrly,df200507hrly,df200509hrly,df200511hrly)
df2005hpd<-rbind(df200503hpd,df200507hpd,df200509hpd,df200511hpd)
rm(df200503hpd,df200507hpd,df200509hpd,df200511hpd)
```
## Structure of HPD and Hourly
```{r}
str(df2005hpd)
str(df2005hrly)
```
## Missing Values Check and Imputation
```{r}
sort(colSums(is.na(df2005hrly))/nrow(df2005hrly)*100, decreasing = T)
sort(colSums(is.na(df2005hpd))/nrow(df2005hpd)*100, decreasing = T)
```
## Let's use Central imputation from DmWR package to impute missing values
```{r}
df2005hrly <- centralImputation(df2005hrly)
df2005hpd <- centralImputation(df2005hpd)
```
## Check if NA values are filled
```{r}
colSums(is.na(df2005hpd))
colSums(is.na(df2005hrly))
```

## Process hpd data 2005
## Convert and Merge the Timestamp in 2005hpd to Same as train_allst
## Convert ScheduledArrTime to HH:MM format
```{r}
##HPD Test Data
# Convert Time to Numeric and then to HH:MM format
df2005hpd$Time<-as.numeric(df2005hpd$Time)
df2005hpd$Time<-sprintf("%04d",df2005hpd$Time)
format(strptime(df2005hpd$Time, format="%H%M"), format = "%H:%M")
str(df2005hpd)
```
## Merge date and Time in hpd
```{r}
##hpd Test Data
df2005hpd<-unite(df2005hpd,"Timestamp",c("YearMonthDay","Time"),remove = TRUE)
df2005hpd$Timestamp<-ymd_hm(df2005hpd$Timestamp)
str(df2005hpd)
df2005hpdb<-df2005hpd
```

## Aggregate Time into 6-hour-wide intervals for 2005hpd and Test data
```{r}
## Test Data
#############
#
tests<-cut.POSIXt(test_allst$ExpectedArrival, breaks = "6 hours")
tests
testdf<-as.data.frame.Date(tests)
str(testdf)
testdf$tests<-as.POSIXct(testdf$tests)
#testdf$tests<-testdf$tests - 3600
c<-cbind(test_allst,testdf)

test_allst<-c[,-2]  #-->Remove Actual Timestamp
colnames(test_allst)[11]<-"ExpectedArrival"

# HPD data
##########

df2005hpdts<-cut.POSIXt(df2005hpd$Timestamp, breaks = "6 hours")
b<-as.data.frame.Date(df2005hpdts)
str(b)
c<-cbind(df2005hpd,b)
df2005hpd<-c[,-2]  ##-->Remove Actual Time
colnames(df2005hpd)[3]<-"Timestamp"

 
df2005hpd<-aggregate(HourlyPrecip~WeatherStationID+Timestamp,data=df2005hpd,FUN=function(df2005hpd) mean(df2005hpd))
```
## Use merge for joining test_allst with hpd data
```{r}
tsallst_hpd<-merge(x=test_allst,y=df2005hpd,by.x=c("WeatherStationID","ExpectedArrival"),by.y=c("WeatherStationID","Timestamp"),all.x=TRUE)

```

## Remove correlated attributes as had removed in train due to high correlation
* Remove DewPointTemp as it is highly correlated(0.81) to DBT
* Remove WindDirection as it is highly correlated to Windspeed (p-value-2.2e-16)
* Remove Skyconditions as it is highly correlated to Visibility(p-value-2.2e-16)

```{r}
df2005hrly$DewPointTemp=NULL
df2005hrly$WindDirection=NULL
df2005hrly$SkyConditions=NULL
str(df2005hrly)
```

## Convert and Merge the Timestamp in 2005hourly to Same as tsallst_hpd
## Convert ScheduledArrTime to HH:MM format
```{r}
##Hourly Test Data
# Convert Time to Numeric and then to HH:MM format

df2005hrlyb<-df2005hrly
df2005hrly$Time<-as.numeric(df2005hrly$Time)
df2005hrly$Time<-sprintf("%04d",df2005hrly$Time)
format(strptime(df2005hrly$Time, format="%H%M"), format = "%H:%M")
str(df2005hrly)
```

## Merge date and Time in hourly test
```{r}
##hourly Test Data
df2005hrlyb<-df2005hrly
df2005hrly<-unite(df2005hrly,"Timestamp",c("YearMonthDay","Time"),remove = TRUE)
df2005hrly$Timestamp<-ymd_hm(df2005hrly$Timestamp)
str(df2005hrly)
```

## Aggregate Time into 6-hour-wide intervals for 2005hourly data
```{r}
# Hourly data
#############
# Aggregate each Timestamp into 6 hour window
df2005hrlyts<-cut.POSIXt(df2005hrly$Timestamp, breaks = "6 hours")
b<-as.data.frame.Date(df2005hrlyts)
str(b)
c<-cbind(df2005hrly,b)
df2005hrly<-c[,-2]
colnames(df2005hrly)[8]<-"Timestamp"


df2005hrly<-aggregate(cbind(Visibility,DBT,RelativeHumidityPercent,WindSpeed,WindGustValue,StationPressure)~WeatherStationID+Timestamp,data=df2005hrly,FUN=function(df2005hrly) mean(df2005hrly))
```

## Merge hourly data with tsallst_hpd
```{r}
tsallsthpd_hrly<-merge(x=tsallst_hpd,y=df2005hrly,by.x=c("WeatherStationID","ExpectedArrival"),by.y=c("WeatherStationID","Timestamp"),all.x=TRUE)
```
## Check for NA values in the final tsallsthpd_hrly dataframe
## Missing Values Check and Imputation
```{r}
sort(colSums(is.na(tsallsthpd_hrly))/nrow(tsallsthpd_hrly)*100, decreasing = T)
```
## Let's use Central imputation from DmWR package to impute missing values
```{r}
tsallsthpd_hrly <- centralImputation(tsallsthpd_hrly)
```
## Check if NA values are filled
```{r}
colSums(is.na(tsallsthpd_hrly))
```
##   #######################End of Weather Data Processing#################

## Split Categorical and Numerical Attributes
```{r}
cat_attr<-c("FlightNumber","Origin","Destination","FlightDelayStatus","TimeZone","ExpectedArrival")
num_attr<-setdiff(colnames(train_data),cat_attr)
num_attr
```
## Check Correlation Plot
```{r}
num_attr_df<-data.frame(train_data[,num_attr])
cor_num_attr_df<-cor(num_attr_df)
#corrplot(cor_num_attr_df, method = "square")
corrplot(cor_num_attr_df, method = "number")
```

* Remove GroundHeight as it is highly negatively correlated to StationPressure(-0.96)
* Remove TimeZone as it is highly correlated to ExpectedArrival
```{r}
train_data$GroundHeight=NULL
test_data$GroundHeight=NULL
train_data$TimeZone=NULL
test_data$TimeZone=NULL
colnames(train_data)
```
## Backups of Data and Removal of attributes as needed
```{r}
## Backups of main data
trn_data<-train_data
tst_data<-test_data

#train_data<-trn_data
#test_data<-tst_data

train_data<-trallsthpd_hrly
test_data<-tsallsthpd_hrly
FlightNumber<-tst_data$FlightNumber
```
# Remove Origin and Destination as the data required is captured in WeatherStationID and ScheduledTravelTime
```{r}
delete <- c("Origin","Destination","FlightNumber")
train_data<-train_data[,!(colnames(train_data) %in% delete),drop=FALSE]
test_data<-test_data[,!(colnames(test_data) %in% delete),drop=FALSE]
```
## Bivariate Analysis
* DBT and RelativeHumidityPercent(Slight negative correlation)
```{r}
ggplot(train_data,aes(x=DBT,
           y=RelativeHumidityPercent)) + 
geom_point(size=2)+ggtitle("DBT VS RelativeHumidityPercent")
```
# ############################### Basic Model Building #############################
## Train-Validation split
```{r}
set.seed(715)
train_rows <- createDataPartition(train_data$FlightDelayStatus, p = 0.75, list = F )
train_data <- train_data[train_rows, ]
validation_data <- train_data[-train_rows, ]
```
## Check for the proportion of Target distribution of classes
```{r}
table(train_data$FlightDelayStatus)
table(validation_data$FlightDelayStatus)
```
* Confirm equal distribution of Status in train & validation.Imbalanced Data;Have to balance in further processing
```{r}
prop.table(table(train_data$FlightDelayStatus))
prop.table(table(validation_data$FlightDelayStatus))
#class(train_data$FlightDelayStatus)
``` 
## Make equal distribution of Target using SMOTE-->Do later
```{r}
#
```

# Build C5.0 Decision Trees
```{r}
#class(train_data$ExpectedArrival)-->POSIXt
#trndbfex<-train_data
#tstbfex<-test_data
#valbfex<-validation_data

train_data$ExpectedArrival=as.numeric(train_data$ExpectedArrival)
validation_data$ExpectedArrival=as.numeric(validation_data$ExpectedArrival)
test_data$ExpectedArrival=as.numeric(test_data$ExpectedArrival)
c5_tree <- C5.0(FlightDelayStatus ~ ., data = train_data)
summary(c5_tree)
plot(c5_tree)
```
##Prediction on Validation data
```{r}
preds_val<-predict(c5_tree,validation_data)
confusionMatrix(preds_val,validation_data$FlightDelayStatus)
```
## F1 score on Validation data-->90.6%
## Accuracy on Validation Data-->83.89%
```{r}
F1_Score(y_true =validation_data$FlightDelayStatus,y_pred = preds_val)
Accuracy(y_true =validation_data$FlightDelayStatus,y_pred = preds_val)
```
## Test Data Prediction F1 Score-->24.96%
```{r}
pred_test <- predict(c5_tree, test_data, type = "class")
FlightDelayStatus<-as.character(pred_test) 
preds_c50<-cbind(FlightNumber,FlightDelayStatus)
write.csv(preds_c50,"submission_c50.csv",row.names = F)
```
